---
title: "UN report on the relationship between GDP and Life expectancy"
author: "Kevin England"
date: "2023-01-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(knitr)
```

This report was prepared to the attention of the United Nations. It analyzes the relationship between a country's GDP, its life expectancy and CO2 emissions. Our goal is to determine to what degree a country’s economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP.

```{r read-data, include = FALSE}
gapminder_1997 <- read_csv("data/gapminder_1997.csv")
```

```{r gdp-lifeExp-1997-plot, echo=FALSE, fig.width = 4, fig.height = 3, fig.align='center'} 
# make plot of GDP (X) vs. LifeExp (Y)
gapminder_1997%>%
  ggplot(aes(x = gdpPercap, y = lifeExp, color=continent)) +
  geom_point() +
  labs(x = "GDP per Capita", y = "Life Expectancy") +
  scale_color_brewer(palette = 'Set1')
```

```{r basic-country-stats}
nCountries <- 
  gapminder_1997 %>%
  select(country) %>%
  n_distinct()

minGDP <- gapminder_1997%>%
  summarize(round(min(gdpPercap))) %>%
  pull()

maxGDP <- gapminder_1997%>%
  summarize(round(max(gdpPercap)))%>%
  pull()
```

The number of countries in the dataset is `r nCountries`.

The minimum GDP is `r minGDP`.

```{r table}
gapminder_1997%>%
  filter(continent == "Oceania") %>%
  kable()
```



